---
title: "CRISPR_faeces_analysis"
output: html_document
---
```{r}
##Load pacakages
library(phyloseq); packageVersion("phyloseq")
library(Biostrings); packageVersion("Biostrings")
library(ggplot2); packageVersion("ggplot2")
library(iNEXT); packageVersion("iNEXT")
library(ape); packageVersion("ape")
library(vegan); packageVersion("vegan")
library(dplyr); packageVersion("dplyr")
library(cowplot); packageVersion("cowplot")
library(plyr); packageVersion("plyr")
library(sjmisc); packageVersion("sjmisc")
library(data.table); packageVersion("data.table")
library(metacoder);packageVersion("metacoder")
library(hilldiv);packageVersion("hilldiv")
library(car);packageVersion("car")
library(lme4);packageVersion("lme4")
library(RColorBrewer);packageVersion("RColorBrewer")
library("ape")
library(patchwork)
library(ggpubr)
library(rstatix)

#We use the ASV table which resulted from the harsher cutoff in Decontam and processing through LULU for curating the ASV table.
ASVs_deconprev_harsh <- read.csv("Curated_Table_CRISPR_deconprev_harsh.txt", sep = ",")
tax_deconprev_harsh <- read.csv("Curated_Tax_CRISPR_deconprev_harsh.csv", sep = ",")
md <- read.csv("metadata_complete.csv", sep = ",",stringsAsFactors = TRUE)
md <- md %>% filter(!demultiplexing_ID %in% c("Mo-014CR","2ALF002","2MOF001"))
md <- md %>% arrange(Sample_ID)
md$Batch <- as.factor(md$Batch)
rownames(md)<-md$demultiplexing_ID
rownames(tax_deconprev_harsh) <- rownames(ASVs_deconprev_harsh)
colnames(ASVs_deconprev_harsh)<-rownames(md)

Define some pallettes
ugly<- c("#C15444","#C1B244","#26736A","#C17A44","#2D8652","#2E428A","#A83861","#BEE0E9","#D4C57D","#4D3619","#883091","#B3B4E6","#CDCF6E","#6ECFA8","#1F1F1F","#5E5E5E","#0E220B","#CD936A","#62CB8C","#C65353","#C6BA53","#4FB3C4","#8FC757","#6B4424","#EC453C","#F4A590","#D6F7AB","#78F297", "#310E8B","#EA462A")

childish <- c("#B71C1C","#FFC107","#A7FFEB","#FFB74D","#757575","#FF5252","#BA68C8","#8BC34A","#F57F17","#009688","#FF80AB","#304FFE","#455A64","#69F0AE","#FF5722","#9C640C","#F1C40F","#A93226","#76448A","#17A589","#C6BA53","#4FB3C4","#8FC757","#6B4424","#EC453C","#F4A590","#D6F7AB","#78F297", "#310E8B","#EA462A")

#make the phyloseq object
physeq_a<- phyloseq(otu_table(ASVs_deconprev_harsh,taxa_are_rows=TRUE),
                   tax_table(as.matrix(tax_deconprev_harsh)),
                   sample_data(md))

#merge_taxa
physeq = tax_glom(physeq_a, "Genus")

## Create tree
random_tree = rtree(ntaxa(physeq), rooted=TRUE, tip.label=taxa_names(physeq))

#merge the tree to the physeq object
physeq = merge_phyloseq(physeq, random_tree)

#Generate sample based relative abundace (and filter low occuring ASVs)
physeq_filt = prune_samples(sample_sums(physeq) >= 100, physeq)

# relative abundance
physeq_norm = transform_sample_counts(physeq_filt, function(x) x/sum(x))
physeq_norm = phyloseq::filter_taxa(physeq_norm, function(x) var(x) > 0, TRUE)

```
alpha diversity
```{r}
#Genus level alpha diversity
physeq_faeces <- subset_samples(physeq,Sample_type=="faeces")
physeq_faeces <- subset_samples(physeq_faeces,experi=="Tyr")
physeq_faeces <- prune_taxa(taxa_sums(physeq_faeces)>0,physeq_faeces)


ASVtab_faeces <- as.data.frame(physeq_faeces@otu_table)
sample_groups <- as.data.frame(as.matrix(physeq_faeces@sam_data))
sample_groups <- as.data.frame(sample_groups[,c(2,6)])

f_richness <- div_test(ASVtab_faeces,hierarchy = sample_groups,q=0,posthoc = TRUE)
f_Shannon <- div_test(ASVtab_faeces,hierarchy = sample_groups,q=1,posthoc = TRUE)
f_Simpson <- div_test(ASVtab_faeces,hierarchy = sample_groups,q=2,posthoc = TRUE)

#plot the alpha diversities
div_test_plot(f_richness,posthoc=TRUE,threshold=0.05,col=childish)
div_test_plot(f_Shannon,posthoc=TRUE,threshold=0.05,col=childish)
div_test_plot(f_Simpson,posthoc=TRUE,threshold=0.05,col=childish)

#Same as above just unfiltered and not merged into genera(ASV_level)
physeq_faeces <- subset_samples(physeq_a,Sample_type=="faeces")
physeq_faeces <- subset_samples(physeq_faeces,experi=="Tyr")
physeq_faeces <- prune_taxa(taxa_sums(physeq_faeces)>0,physeq_faeces)

ASVtab_faeces <- as.data.frame(physeq_faeces@otu_table)
sample_groups <- as.data.frame(as.matrix(physeq_faeces@sam_data))
sample_groups <- as.data.frame(sample_groups[,c(2,6)])

f_richness <- div_test(ASVtab_faeces,hierarchy = sample_groups,q=0,posthoc = TRUE)
f_Shannon <- div_test(ASVtab_faeces,hierarchy = sample_groups,q=1,posthoc = TRUE)
f_Simpson <- div_test(ASVtab_faeces,hierarchy = sample_groups,q=2,posthoc = TRUE)

#plot the alpha diversities
div_test_plot(f_richness,posthoc=TRUE,threshold=0.05,col=childish)
div_test_plot(f_Shannon,posthoc=TRUE,threshold=0.05,col=childish)
div_test_plot(f_Simpson,posthoc=TRUE,threshold=0.05,col=childish)

```
Relative Abundances
```{r}

#filter out the faeces
physeq_faeces <- subset_samples(physeq_norm,Sample_type=="faeces")
physeq_faeces <- subset_samples(physeq_faeces,experi=="Tyr")
physeq_faeces <- prune_taxa(taxa_sums(physeq_faeces)>0,physeq_faeces)

#create physeq objects for each phenotype just in case its needed
physeq_wt <- subset_samples(physeq_faeces,Fish_type=="WT")
physeq_al <- subset_samples(physeq_faeces,Fish_type=="Albino")
physeq_mo <- subset_samples(physeq_faeces,Fish_type=="Mosaic")

#Lets get all the genera found in the faeces samples

genera <- as.data.frame(physeq_faeces@tax_table)
genera <- subset(genera, select=-Kingdom)

abundance_full <- as.data.frame(taxa_sums(physeq_faeces)/nsamples(physeq_faeces))
abundance_WT  <- as.data.frame(taxa_sums(physeq_wt)/nsamples(physeq_wt))
abundance_AL  <- as.data.frame(taxa_sums(physeq_al)/nsamples(physeq_al))
abundance_Mo  <- as.data.frame(taxa_sums(physeq_mo)/nsamples(physeq_mo))

generalist <- cbind(genera,abundance_full,abundance_WT,abundance_AL,abundance_Mo)
generalist <- generalist %>% 

write.table(generalist,"generalist.tsv",sep="\t", quote=F, col.names=NA)

#plot time
par(mar = c(10, 4, 4, 2) + 0.1) # make more room on bottom margin
N <- 30

#tiff("Rank_Barplot_30_faeces_postdecon_harsh.tiff", units="in", width=12, height=6, res=300)
barplot(sort(taxa_sums(physeq_faeces), TRUE)[1:N]/nsamples(physeq_faeces), las=2,col="#9C640C")+ title(main="Cumulative fractional abundance of top 30 ASV's in the faeces samples",ylab = "Fractional abundance") 
#dev.off()

#Cumulative relative abundance accross fecal samples
#Cetobacterium
sort(taxa_sums(physeq_faeces), TRUE)[1]/nsamples(physeq_faeces)
#Aeromonas
sort(taxa_sums(physeq_faeces), TRUE)[2]/nsamples(physeq_faeces)

#selecting top30 genera
TopNOTUs = names(sort(taxa_sums(physeq_faeces), TRUE)[1:20])
physeq_30_f = prune_taxa(TopNOTUs, physeq_faeces)

#Plot based on class
plot_class_ab_f = plot_bar(physeq_faeces, "demultiplexing_ID", fill = "Class") + 
  geom_bar(aes(color=Class, fill=Class), stat="identity", position="stack") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90),strip.text.x = element_text(colour = "black",face = "bold"),strip.background = element_blank()) +
  labs(x = "Samples") +
  labs(y = "Relative Abundance") + 
  scale_fill_manual(values=ugly) +
  scale_color_manual(values=ugly) 

plot_class_ab_f = plot_class_ab_f + ggforce::facet_row(~Fish_type+Sample_type, scales = "free_x", space = "free",) + guides(fill=guide_legend(ncol=1))

#plot based on genus
plot_genus_ab_f = plot_bar(physeq_30_f, "demultiplexing_ID", fill = "Genus") + 
  geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") +
  theme_minimal() +
 theme(axis.text.x = element_blank(),strip.background = element_rect(colour = "gray"), strip.text = element_text(face = "bold",size=20),panel.background = element_rect(color = "black"),panel.grid = element_blank(),title = element_text(size = 20),axis.text.y = element_text(size=10),legend.text = element_text(size=20,face="italic")) +
  labs(title = "Individual faeces samples",x="") +
  labs(y = "Relative Abundance") + 
  scale_fill_manual(values=childish) +
  scale_color_manual(values=childish) +
  scale_y_continuous(expand = c(0,0))

plot_genus_ab_f = plot_genus_ab_f + ggforce::facet_row(~Fish_type, scales = "free_x", space = "free",strip.position = "bottom") + guides(fill=guide_legend(ncol=1)) 


#Make the merged plot

physeq_faeces_final <- subset_samples(physeq_filt, Sample_type =="faeces")
physeq_faeces_final <- subset_samples(physeq_faeces_final, experi == "Tyr")
#Make new variable by which to merge the samples
btype <- as.character(get_variable(physeq_faeces_final,"Fish_type"))
sample_data(physeq_faeces_final)$sample_batch<- mapply(paste0,btype,collapse="_")

#merge by new variable
physeq_faeces_final <- merge_samples(physeq_faeces_final, "sample_batch")
#clc realtive abundnce and normailise
physeq_faeces_final<- transform_sample_counts(physeq_faeces_final, function(x) x/sum(x))
physeq_faeces_final = phyloseq::filter_taxa(physeq_faeces_final, function(x) var(x) > 0, TRUE)
sample_names(physeq_faeces_final) 

#plot
physeq_30_fm = prune_taxa(TopNOTUs, physeq_faeces_final)

physeq_faeces_final_Genus = plot_bar(physeq_30_fm ,fill = "Genus") + 
  geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") +
  theme_minimal() +
  theme(axis.text.x = element_text(colour = "black",face = "bold",size = 20,vjust = 8), strip.background = element_rect(colour = "gray"),panel.background = element_rect(color = "black"),panel.grid = element_blank(),title = element_text(size=20),axis.text.y = element_text(size=10),axis.title.x = element_blank(),legend.text = element_text(size=20,face="italic")) +
  labs(x="",title = "Merged faeces samples") +
  labs(y = "Relative Abundance") + 
  scale_fill_manual(values=childish) +
  scale_color_manual(values=childish) +
  scale_y_continuous(expand = c(0,0))


faeces_abundance <- physeq_faeces_final_Genus + plot_genus_ab_f + plot_layout(guides="collect",widths = c(1,2)) + plot_annotation(tag_levels = "A")
```
metacoder
```{r}
#preprocessing to make the metacoder plot
metacoder <- parse_phyloseq(physeq_faeces)
metacoder$data$tax_abund <- calc_taxon_abund(metacoder, data = "otu_table") # fine
metacoder$data$tax_occ <- calc_n_samples(metacoder, "tax_abund", groups = "Fish_type")
metacoder$data$diff_table <- compare_groups(metacoder, data = "tax_abund", cols = metacoder$data$sample_data$demultiplexing_ID,
                                            groups = metacoder$data$sample_data$Fish_type)
metacoder$data$diff_table$adjusted_p_value <- p.adjust(metacoder$data$diff_table$wilcox_p_value,
                                                       method = "BH")

View(metacoder$data$diff_table)
metacoder$data$diff_table$log2_median_ratio[metacoder$data$diff_table$wilcox_p_value > 0.05] <- 0 # added

#metacoder plot time
metacoder.plot <- heat_tree_matrix(metacoder,
                                   data = "diff_table",
                            node_size = n_obs,
                            node_label = taxon_names,
                            node_color = log2_median_ratio,
                            node_color_range = diverging_palette(),
                            node_color_trans = "linear",
                            node_color_interval = c(-3, 3),
                            edge_color_interval = c(-3, 3),
                            overlap_avoidance = 5,
                            layout = "fruchterman-reingold", # The primary layout algorithm
                            initial_layout = "reingold-tilford", # The layout algorithm that initializes node locations
                            node_size_axis_label = "Number of ASVs",
                            node_color_axis_label = "Log2 ratio median proportions") 
pdf("Compare_faeces_types.pdf", width = 20, height = 20)
metacoder.plot
dev.off()

#Show the taxon_IDs of the genera which differentiated significantly among the pigmentation phenotypes
significant_taxa<- metacoder$data$diff_table %>%
  mutate(taxon_names = taxon_names(metacoder)[taxon_id], 
         taxon_ranks = taxon_ranks(metacoder)[taxon_id]) %>%
  select(taxon_id, 
         taxon_names,
         taxon_ranks,
         everything()) %>% # put names in front (optional)
  filter(taxon_ranks == "Genus") %>%
  arrange(desc(abs(median_diff)))
          
#plotting only the taxa that are differrent between the Albino and WT
          
          physeq_two <- subset_samples(physeq_faeces,Fish_type!="Albino")
          physeq_two <- prune_taxa(taxa_sums(physeq_two)>0,physeq_two)
        
          metacoder <- parse_phyloseq(physeq_two)
metacoder$data$tax_abund <- calc_taxon_abund(metacoder, data = "otu_table") # fine
metacoder$data$tax_occ <- calc_n_samples(metacoder, "tax_abund", groups = "Fish_type")
metacoder$data$diff_table <- compare_groups(metacoder, data = "tax_abund", cols = metacoder$data$sample_data$demultiplexing_ID,
                                            groups = metacoder$data$sample_data$Fish_type)
metacoder$data$diff_table$adjusted_p_value <- p.adjust(metacoder$data$diff_table$wilcox_p_value,
                                                       method = "fdr")

View(metacoder$data$diff_table)
metacoder$data$diff_table$log2_median_ratio[metacoder$data$diff_table$wilcox_p_value > 0.05] <- 0 # added

metacoder.plot <- heat_tree(metacoder,
                            node_size = n_obs,
                            node_label = taxon_names,
                            node_color = log2_median_ratio,
                            node_color_range = diverging_palette(),
                            node_color_trans = "linear",
                            node_color_interval = c(-3, 3),
                            edge_color_interval = c(-3, 3),
                            overlap_avoidance = 5,
                            layout = "fruchterman-reingold", # The primary layout algorithm
                            initial_layout = "reingold-tilford", # The layout algorithm that initializes node locations
                            node_size_axis_label = "Number of ASVs",
                            node_color_axis_label = "Log2 ratio median proportions") 

jpeg("metacoder_MO_WT.jpeg", units="in", width=15, height=20, res=300)
#pdf("Compare_faeces_Wt_al.pdf", width = 20, height = 20)
metacoder.plot # for ewos green more abundant in gut, no enrichment of mycoplasma so it must also be abundant in the feed # bremeia is more abudant in the gut
dev.off()

#Repeat the above for each individual comparison
```

```{r}
#Boxplots (including stats) showing the significant differences of the significantly different genera which constitute more than 1% of the cumulative relative abundance. These are Cetobacterium, Aeromonas, Vibrio and Pseudomonas.

Fish_types <- as.data.frame(as.matrix(physeq_faeces@sam_data$Fish_type))
#Cetobacterium abundance
Cetobacterium <- physeq_faeces@otu_table["ASV_1"]
Cetobacterium <- t(Cetobacterium)
#Aeromonas
Aeromonas<- physeq_faeces@otu_table["ASV_3"]
Aeromonas <- t(Aeromonas)
#Vibrio abundance
Vibrio<- physeq_faeces@otu_table["ASV_24"]
Vibrio <- t(Vibrio)
#Pseduomonas abundance
Pseduomonas <- physeq_faeces@otu_table["ASV_13"]
Pseduomonas <- t(Pseduomonas)

#plesio
final <- cbind(Cetobacterium,Aeromonas,Vibrio,Pseduomonas,Fish_types)

#Cetobacterium
#Stat_tests for including in the plots. 
cet <- compare_means(ASV_1~V1,data=final,method = "wilcox.test",p.adjust.method = "BH",hide.ns=TRUE) %>% add_significance("p.adj")  %>% add_xy_position(formula=ASV_1~V1,data=final,dodge = 0.8)
#plot
Ceto <- ggboxplot(final, x="V1",y="ASV_1",fill = childish[1:3])+labs(x="",y="Relative abundance") + stat_pvalue_manual(cet,label = "p.adj.signif",hide.ns = TRUE) + labs(title = "Cetobacterium") + theme(title = element_text(face = "italic"),axis.title.y = element_text(face = "plain"))
#Aeromonas
Aer <- compare_means(ASV_3~V1,data=final,method = "wilcox.test",p.adjust.method = "BH") %>% add_significance("p.adj")  %>% add_xy_position(formula=ASV_3~V1,data=final,dodge = 0.8)

Aero <- ggboxplot(final, x="V1",y="ASV_3",fill = childish[1:3])+labs(x="",y="Relative abundance") + stat_pvalue_manual(Aer,label = "p.adj.signif",,hide.ns = TRUE) + labs(title = "Aeromonas") + theme(title = element_text(face = "italic"),axis.title.y = element_text(face = "plain"))

#Vibrio

Vi <- compare_means(ASV_24~V1,data=final,method = "wilcox.test",p.adjust.method = "BH") %>% add_significance("p.adj")  %>% add_xy_position(formula=ASV_24~V1,data=final,dodge = 0.8)

Vib <- ggboxplot(final, x="V1",y="ASV_24",fill = childish[1:3])+labs(x="",y="Relative abundance") + stat_pvalue_manual(Vi,label = "p.adj.signif",,hide.ns = TRUE) + labs(title = "Vibrio") + theme(title = element_text(face = "italic"),axis.title.y = element_text(face = "plain"))

#Pseudomonas

pse <- compare_means(ASV_13~V1,data=final,method = "wilcox.test",p.adjust.method = "BH") %>% add_significance("p.adj")  %>% add_xy_position(formula=ASV_13~V1,data=final,dodge = 0.8)

Ps <- ggboxplot(final, x="V1",y="ASV_13",fill = childish[1:3])+labs(x="",y="Relative abundance") + stat_pvalue_manual(pse,label = "p.adj.signif",,hide.ns = TRUE) + labs(title = "Pseudomonas")  + theme(title = element_text(face = "italic"),axis.title.y = element_text(face = "plain"))

taxas <- (Ceto + Aero) / (Vib + Ps) + plot_annotation(tag_levels = "A",title="Significant taxa") & theme(title = element_text(size=20))
```
Beta diversities (ordinations)
```{r}

#_Weighted and unweighted UniFrac ordinations
ordipoo_wuni <- ordinate(physeq_faeces,"PCoA","unifrac",weighted=TRUE)
ordipoo_ununi <- ordinate(physeq_faeces,"PCoA","unifrac",weighted=FALSE)

#plot the ordinations
#weighted UniFrac
p1 = plot_ordination(physeq_faeces, ordipoo_wuni, type="Samples",
                     color="Fish_type", shape = "Batch",
                     title="PCoA")
pordi_poo_wuni = p1+ geom_point(size = 5) + 
  theme_minimal() + theme(axis.title = element_text(size = 15)) + 
  labs(subtitle = "Weighted UniFrac") +
    scale_fill_manual(values=childish) + 
  scale_color_manual(values=childish,name="Pigmentation phenotype")+
  scale_shape_discrete(name="Sequencing run")
  
#unweighted UniFrac
p1 = plot_ordination(physeq_faeces, ordipoo_ununi, type="Samples",
                     color="Fish_type", shape = "Batch",
                     title="PCoA")
pordi_poo_uni = p1+ geom_point(size = 5) + 
  theme_minimal() + theme(axis.title = element_text(size = 15)) + 
  labs(subtitle = "Unweighted UniFrac") +
    scale_fill_manual(values=childish) + 
  scale_color_manual(values=childish,name="Pigmentation phenotype")+
  scale_shape_discrete(name="Sequencing run")

#PERMANOVA using Bray-Curtis
adonis(t(as.data.frame(physeq_faeces@otu_table))~Fish_type*Batch,data=as.data.frame(as.matrix(physeq_faeces@sam_data)),method="bray")

